---
title: "STA130 -- Fall 2021"
author: "[ADD YOUR NAME HERE]"
subtitle: Module 2 Problem Set 
output:
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(openintro)
```


# Instructions

## How do I hand in these problems for the Thursday deadline ?

Your complete .Rmd file that you create for these practice problems AND the resulting pdf (i.e., the one you 'Knit to PDF' from your .Rmd file) must be uploaded into a Quercus assignment (link: [https://q.utoronto.ca/courses/206597/assignments/707456](https://q.utoronto.ca/courses/206597/assignments/707456)) by 11:59 a.m. ET, on Thursday, Sep 23. Late problem sets are not accepted.



# Problem set grading

There are two parts to your problem set. One is largely R-based with short written answers and the other is more focused on writing. We recommend you use a word processing software like Microsoft Word to check for grammar errors in your written work. Note: there can be issues copying from Word to R Markdown so it may be easier to write in this file first and then copy the text to Word. Then you can make any changes flagged in Word directly in this file.


# Part 1


## [Question 1] The `games.csv` file contains data for a sample of video games, available through Steam, with under 1000 players on at a time,

You will be working with the following variables featured in the csv file:
 - `X1` (double) an identification number
 - `gamename`	(character)	Name of video games
 - `year`	(double)	Year of measure
 - `month`	(character)	Month of measure
 - `avg`	(double)	Average number of players at the same time
 - `gain`	(double)	Gain (or loss) Difference in average compared to the previous month (NA = 1st month)
 - `peak`	(double)	Highest number of players at the same time
 - `avg_peak_perc`	(character)	Share of the average in the maximum value (avg / peak) in %

Source: [https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-16/readme.md](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-16/readme.md)

```{r, message=FALSE}
library(tidyverse) # Load the tidyverse package so it is available to use 
games <- read_csv("games.csv")
```

### (a) Use the `glimpse()` function to view properties of the `games` data set. How many rows and columns are there? How many observations does it include? How many variables are measured for each observation? 

```{r}


```

*<Type your answer to the written question here (note that R code does not run outside of the grey R code chunks).>*



### (b) Create 3 histograms to explore the distribution of average number of players at the same time: (i) one with 2 bins, (ii) one with 10 bins, and (iii) one with 80 bins; make sure to specify meaningful axis labels where appropriate. Which of these histograms is most appropriate to describe the distribution of the average number of players at the same time? Why? Write a few sentences describing the distribution based on the histogram you chose as most appropriate.

```{r}
 


```


*<Type your answer to the written question here (note that R code does not run outside of the grey R code chunks).>*


### (c) Construct a plot to visualize the distribution of a categorical variable from the games data and describe the distribution in 1-2 sentences; make sure to specify meaningful axis labels where appropriate. Hint: If you choose a categorical variable with many different categories, you may find it useful to use `coord_flip()` to flip the bars horizontally and/or change the options in the R code chunk to make the plot large (ex: {r, fig.height=15, fig.width=5}).

```{r}

```


*<Type your answer to the written question here (note that R code does not run outside of the grey R code chunks).>*




### (d) Construct a set of three boxplots showing visual summaries of the distribution of highest number of players on at a time (`peak`) for each month (`month`); make sure to specify meaningful axis labels where appropriate. Write 3-4 sentences comparing these distributions.

```{r}

```


*<Type your answer to the written question here (note that R code does not run outside of the grey R code chunks).>*





## [Question 2] The `births` data set is part of the `openintro` package. It consists of random sample of 100 births for babies in North Carolina where the mother was not a smoker and another 50 where the mother was a smoker.   Type `?births` in the R console for more information about the data and to see the definition of each variable. The code below loads the required libraries for this question and provides a glimpse of the `births` data frame.
  
```{r, include=FALSE}
library(tidyverse)
library(openintro)
```

```{r}
glimpse(births)
```
  
  
### (a) Choose two categorical variables and plot the distribution of each one (in seperate plots). Identify whether each of these variables is a nominal or ordinal categorical variable. Write one or two sentences interpreting each plot.  

```{r}

```

*<Type your answer to the written question here (note that R code does not run outside of the grey R code chunks).>*



### (b) Choose a quantitative variable and plot its distribution.  Identify whether the variable you selected is continuous or discrete, and write 2-3 sentences describing the distribution.

```{r}



```

*<Type your answer to the written question here (note that R code does not run outside of the grey R code chunks).>*


### (c) Construct a plot that shows the relationship between birth weight (`weight`) and mother's smoking status (`smoke`); make sure to specify meaningful axis labels where appropriate.  

```{r}



```



\newpage

# Part 2

## Writing prompt:  

Pretend that you are on the phone with your friend, and you want to share some of the cool new data visualization techniques that you have been learning in STA130. Pick one of the above graphs and prepare a small paragraph on how you would describe the graph to your friend (keeping in mind that they cannot see the graph). Make sure you include at least 2 words/phrases from your vocabulary list. It is also important to keep in mind that the person you are talking to has not taken STA130, therefore they will not be as familiar with the statistical vocabulary as you are. Therefore, make sure the explain any terms you use in plain language.  

 

### When describing a figure, it is important to:  

- Describe the data source  
- State the type of graph  
- Identify what are on the x- and y-axis (if appropriate)  
- Describe the distribution  
- Make note of potential outliers  

 
### Vocabulary List  

- Where are the data centered (approximate values if available)  
- How much spread (relative to what?)  
- Shape: symmetric, left-skewed, right-skewed  
- The tails of the distribution (heavy-tailed or thin-tailed)  
- Modes: where, how many, unimodal, bimodal, multimodal, uniform  
- Outliers, extreme values  
- Frequency (which category occurred the most or least often; data concentrated near a particular value or category)  
- Mean, median, mode  
- Standard deviation, interquartile range  

 

### Some general reminders  

- Try to not send more than 20 minutes on the prompt.  
- Aim for more than 200 words but less than 500 words.  
- Use full sentences. 
- Grammar is *not* the main focus of the assessment, but it is important that you communicate in a clear and professional manner. I.e., no slang or emojis should appear. 
- Be specific. A good principle when responding to a writing prompt in STA130 is to assume that your audience is not aware of the subject matter (or in this case has not read the syllabus). Therefore, in this case, you need to properly communicate what the objective is putting it in your own words (paraphrasing). You CANNOT just copy the learning objective and put it in quotation marks.  